---
title: "Austrailian wild fires"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(here)
library(plyr)
library(dplyr)
library(ggplot2)
library(rgdal)
library(tmap)
library(ggmap)
library(dplyr)
library(ggspatial)
library(broom)
library(tidyverse)
library(readxl)
library(purrr)
library(sf)
library(ozmaps)
library(mapview)
library(plotly)
library(lubridate)
library(shiny)
library(leaflet)
library(DT)
library(tigris)
library(viridis)
library(MAP)
library(ozmaps)
library(ggrepel)

# Loading the data
nasa_fire <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-07/MODIS_C6_Australia_and_New_Zealand_7d.csv')
rainfall <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-07/rainfall.csv')
temperature <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-07/temperature.csv')
url <- "http://www.rfs.nsw.gov.au/feeds/majorIncidents.json"

name_df <- tribble(
  ~station_code, ~city_name, ~lat, ~long, ~station_name,
  "009151", "Perth", -31.96, 115.79, "Subiaco Wastewater Treatment Plant",
  "023011", "Adelaide", -34.92, 138.6, "North Adelaide",
  "040383", "Brisbane", -27.51, 153.05, "Greenslopes Private Hospital",
  "040913", "Brisbane", -27.48, 153.04, "Brisbane",
  "066062", "Sydney", -33.86, 151.21, "Observatory Hill",
  "070351", "Canberra", -35.31, 149.2, "Canberra Airport",
  "086232", "Melbourne", -37.83, 144.98, "Melbourne Botanical Gardens"
)
```


Sidebar {.sidebar}
======================================================================
Use the _Explore rainfall_ tab to explore the the rainfall by the major cities in Australia. The blue line on the time series plot shows long term trend of monthly rainfall. Click and drag on the time series plot to highlight the date and the corresponding monthly average rainfall. 

Use the _Rain fall_ tab to explore the the temperature by the major cities in Australia. The red and yellow line shows long term trend of minimum and maximum temperature. Search the data table to view the date and corresponding daily temperature.

Use the _Maps_ tab to explore wild fire in the week of December 29th, 2019 in Australia.The red dots indicates the fire, with the blue dots indicating the weather stations. click on the map to highlight the weather station and view fire around the station.


Application Authors: Alison Barbee, Kate Bazany, Chukwudi Keke


Data source: (http://www.bom.gov.au/?ref=logo) Bureau of Metrology, Australia



Rain fall
=====================================================================

Column
----------------------------------------------------------------------

### Rain fall plot

```{r}
clean_rain_df_3 <- rainfall %>%
  filter(year > 2009) %>%
  mutate(year_month = paste(year, month, sep = "-")) %>%
  mutate(year_month = ym(year_month)) %>%
  group_by(station_code, city_name, lat, long, station_name, year_month) %>%
  dplyr::summarize(monthly_rain = sum(rainfall)) 

rain_plot_monthly <- clean_rain_df_3 %>%
  ggplot() +
  geom_point(aes(x = year_month, y = monthly_rain), 
             color = "blue", alpha = .8, size = .1) + 
  geom_smooth(aes(x = year_month, y = monthly_rain), color = "black") + facet_wrap(vars(city_name)) +
  theme_classic() + ylab("Total Monthly Rainfall (cm)") + 
  xlab("Date (year)")

ggplotly(rain_plot_monthly)
```

Temperature
============================================================================

Column {data-width=500}
----------------------------------------------------------------------

### Temperature plots

```{r}
clean_temp_df <- temperature %>%
group_by(city_name, temp_type, site_name, date) %>%
dplyr::summarize(average_temp = mean(temperature, na.rm = TRUE)) %>%
pivot_wider(names_from = temp_type, values_from = average_temp)

clean_temp_df_all <- temperature %>% 
  pivot_wider(names_from = temp_type, values_from = temperature)
clean_temp_df_all$year <- substr(clean_temp_df_all$date, 1, 4)
clean_temp_df_10 <- clean_temp_df_all %>%
  filter(year > 2009)

temp_plot_10 <- clean_temp_df_10 %>%
  ggplot() +
  geom_point(aes(x = date, y = max), size = .1, alpha = .05, color = "red") +
  geom_point(aes(x = date, y = min), size = .1, alpha = .05, color = "orange") +
  geom_smooth(aes(date, max), color = "red") +
  geom_smooth(aes(date, min), color = "orange") +
  facet_wrap(vars(city_name)) +
  theme_classic() + ylab("Max and Min Temp (C)") + xlab("Date (year)")

temp_plot_10
```

Column {data-width=350}
----------------------------------------------------------------------

### Temperature data table 
```{r}
clean_temp_df %>%
  
  rename(`site name` = site_name,
         `city name` = city_name) %>%
  
datatable()
```


Maps
============================================================================

Column {data-width=350}
----------------------------------------------------------------------
```{r}
aus_stations <- name_df %>%
  dplyr::select(station_code, city_name, lat, long, station_name)

aus_stations <- st_as_sf(aus_stations, coords = c("long", "lat")) %>%
  st_set_crs(4283)

# Creates the datasets that will be used for creating the ggplot and leaflet maps

aus_fires1 <- nasa_fire %>%
  dplyr::select(latitude, longitude, brightness, acq_date)

aus_fires <- st_as_sf(aus_fires1, coords = c("longitude", "latitude")) %>%
  st_set_crs(4283)

oz_states <- ozmaps::ozmap_states

ggplot() + 
  geom_sf(data = oz_states) + 
  geom_sf(data = aus_fires, colour = "red", shape = 17) +
  geom_sf(data = aus_stations, colour = "darkgreen", fill = "springgreen", 
          size = 3, shape = 22) + 
  ggtitle("Australia Fires During the Week of Dec. 29th, 2019 - Jan. 5th, 2020") +
  xlab("Longitude") + ylab("Latitiude") +
  coord_sf() +
  scale_colour_manual(values = c("fire" = "red", "stations" = "darkgreen"), name = NULL,
                      guide = guide_legend(override.aes = list(shape = c(17, 22)))) +
  theme_bw()
```

Column {data-width=650}
----------------------------------------------------------------------

### Weather stations

```{r}
#This code will generate a leaflet map of Australia which includes data points of all fires within a week as well as the weather stations

leaflet() %>% 
  addTiles() %>% 
  setView(130, -25, zoom = 4) %>%
  addMarkers(115.79, -31.96, popup = "Subiaco Wastewater Treatment Plant, Perth") %>%
  addMarkers(138.6, -34.92, popup = "North Adelaide, Adelaide") %>%
  addMarkers(153.05, -27.51, popup = "Greenslopes Private Hospital, Brisbane") %>%
  addMarkers(153.04, -27.48, popup = "Brisbane Station, Brisbane") %>%
  addMarkers(151.21, -33.86, popup = "Observatory Hill, Sydney") %>%
  addMarkers(149.2, -35.31, popup = "Canberra Airport, Canberra") %>%
  addMarkers(144.98, -37.83, popup = "Melbourne Botanical Gardens, Melbourne") %>%
  addCircles(data = aus_fires1, color = "red", popup = "fire",
             fillOpacity = 0.25, radius = 30000) 
  
```