---
title: "Temperature and Precipitation Leading up to the Devistating Fire Events of 2019-2020"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(here)
library(plyr)
library(dplyr)
library(ggplot2)
library(rgdal)
library(tmap)
library(ggmap)
library(dplyr)
library(ggspatial)
library(broom)
library(tidyverse)
library(readxl)
library(purrr)
library(sf)
library(ozmaps)
library(mapview)
library(plotly)
library(lubridate)
library(shiny)
library(leaflet)
library(DT)
library(tigris)
library(viridis)
library(MAP)
library(ggrepel)

# Loading the data
nasa_fire <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-07/MODIS_C6_Australia_and_New_Zealand_7d.csv')
rainfall <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-07/rainfall.csv')
temperature <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-07/temperature.csv')
url <- "http://www.rfs.nsw.gov.au/feeds/majorIncidents.json"

name_df <- tribble(
  ~station_code, ~city_name, ~lat, ~long, ~station_name,
  "009151", "Perth", -31.96, 115.79, "Subiaco Wastewater Treatment Plant",
  "023011", "Adelaide", -34.92, 138.6, "North Adelaide",
  "040383", "Brisbane", -27.51, 153.05, "Greenslopes Private Hospital",
  "040913", "Brisbane", -27.48, 153.04, "Brisbane",
  "066062", "Sydney", -33.86, 151.21, "Observatory Hill",
  "070351", "Canberra", -35.31, 149.2, "Canberra Airport",
  "086232", "Melbourne", -37.83, 144.98, "Melbourne Botanical Gardens"
)
```


Sidebar {.sidebar}
======================================================================
Australia had significant increases in temperatures and a prolonged drought which led to the devastating fires in December of 2019 that effected millions of the country's residents. The data was pulled from NASA records during the week of December 29th, 2019, to January 5th, 2020. 

Use the _Rainfall_ tab to explore the total monthly rainfall (in centimeters) measured at six field stations across Australia from 2010 to 2020. 

Use the _Temperature_ tab to explore the daily maxiumum and minimum temperatures in degrees Celsius recorded in seven field stations across Australia from 2010 to 2020. 

Use the _Maps_ tab to explore a map of Australia showing fire data collected from NASA records during the week of December 29th, 2019, to January 5th, 2020. 


Application Authors: Alison Barbee, Kate Bazany, Chukwudi Keke


Data source: (http://www.bom.gov.au/?ref=logo) Bureau of Meteorology, Australia



Rainfall
=====================================================================

Column
----------------------------------------------------------------------

### Total monthly rainfall gradually decreased over the ten years leading up to the fire. The blue points represent the total rainfall in centimeters collected at each field station. The line on the time series plots show the general trend of total monthly rainfall over the ten years leading up to the fire events. These rainfall plots indicate that there were slightly decreases in monthly total rainfall in the years leading up to the fire events in the last week of 2019 and first week of 2020. Click and drag on the time series plot to highlight the date and the corresponding monthly average rainfall. 

```{r}
clean_rain_df_3 <- rainfall %>%
  filter(year > 2009) %>%
  mutate(year_month = paste(year, month, sep = "-")) %>%
  mutate(year_month = ym(year_month)) %>%
  group_by(station_code, city_name, lat, long, station_name, year_month) %>%
  dplyr::summarize(monthly_rain = sum(rainfall)) 

rain_plot_monthly <- clean_rain_df_3 %>%
  ggplot() +
  geom_point(aes(x = year_month, y = monthly_rain), 
             color = "blue", alpha = .8, size = .1) + 
  geom_smooth(aes(x = year_month, y = monthly_rain), color = "black") + facet_wrap(vars(city_name)) +
  theme_classic() + ylab("Total Monthly Rainfall (cm)") + 
  xlab("Date (year)")

ggplotly(rain_plot_monthly)
```

Temperature
============================================================================

Column
----------------------------------------------------------------------

### The red points are the daily recorded high temperature and the yellow points are the daily recorded low temperatures. The bold red and yellow lines show the general trend of high and low temperatures over time, made with geom_smooth. This graph shows slight increases in temperature at most stations over the last ten years, as well as a heatwave in the last season leading up to the wildfire events of 2019 and 2020. 

```{r}
clean_temp_df <- temperature %>%
group_by(city_name, temp_type, site_name, date) %>%
dplyr::summarize(average_temp = mean(temperature, na.rm = TRUE)) %>%
pivot_wider(names_from = temp_type, values_from = average_temp)

clean_temp_df_all <- temperature %>% 
  pivot_wider(names_from = temp_type, values_from = temperature)
clean_temp_df_all$year <- substr(clean_temp_df_all$date, 1, 4)
clean_temp_df_10 <- clean_temp_df_all %>%
  filter(year > 2009)

temp_plot_10 <- clean_temp_df_10 %>%
  ggplot() +
  geom_point(aes(x = date, y = max), size = .1, alpha = .05, color = "red") +
  geom_point(aes(x = date, y = min), size = .1, alpha = .05, color = "orange") +
  geom_smooth(aes(date, max), color = "red") +
  geom_smooth(aes(date, min), color = "orange") +
  facet_wrap(vars(city_name)) +
  theme_classic() + ylab("Max and Min Temp (C)") + xlab("Date (year)")

temp_plot_10
```

Map
============================================================================

Column 
----------------------------------------------------------------------

### The red circles on the map represent a 1km fire pixel which may not represent the actual location of the fire since one or more fires can be detected within the 1km pixel. The blue markers represent 7 weather stations around the country that tracked climate data, like temperature and rainfall, in 5 of the 6 states, shown in the Rainfall and Temperature tabs. Click on the map to highlight the weather station and view locations of fires around the station during that week.

```{r}
aus_stations <- name_df %>%
  dplyr::select(station_code, city_name, lat, long, station_name)

aus_stations <- st_as_sf(aus_stations, coords = c("long", "lat")) %>%
  st_set_crs(4283)

# Creates the datasets that will be used for creating the ggplot and leaflet maps

aus_fires1 <- nasa_fire %>%
  dplyr::select(latitude, longitude, brightness, acq_date)

aus_fires <- st_as_sf(aus_fires1, coords = c("longitude", "latitude")) %>%
  st_set_crs(4283)

oz_states <- ozmaps::ozmap_states

#This code will generate a leaflet map of Australia which includes data points of all fires within a week as well as the weather stations

leaflet() %>% 
  addTiles() %>% 
  setView(130, -25, zoom = 4.5) %>%
  addMarkers(115.79, -31.96, popup = "Subiaco Wastewater Treatment Plant, Perth") %>%
  addMarkers(138.6, -34.92, popup = "North Adelaide, Adelaide") %>%
  addMarkers(153.05, -27.51, popup = "Greenslopes Private Hospital, Brisbane") %>%
  addMarkers(153.04, -27.48, popup = "Brisbane Station, Brisbane") %>%
  addMarkers(151.21, -33.86, popup = "Observatory Hill, Sydney") %>%
  addMarkers(149.2, -35.31, popup = "Canberra Airport, Canberra") %>%
  addMarkers(144.98, -37.83, popup = "Melbourne Botanical Gardens, Melbourne") %>%
  addCircles(data = aus_fires1, color = "red", popup = "fire",
             fillOpacity = 0.25, radius = 30000) 
  
```
