---
title: "Austrailian wild fires"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(here)
library(plyr)
library(dplyr)
library(ggplot2)
library(rgdal)
library(tmap)
library(ggmap)
library(dplyr)
library(ggspatial)
library(broom)
library(tidyverse)
library(readxl)
library(purrr)
library(sf)
library(ozmaps)
library(mapview)
library(plotly)
library(lubridate)
library(shiny)
library(leaflet)
library(DT)
# Loading the data
nasa_fire <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-07/MODIS_C6_Australia_and_New_Zealand_7d.csv')
rainfall <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-07/rainfall.csv')
temperature <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-07/temperature.csv')
url <- "http://www.rfs.nsw.gov.au/feeds/majorIncidents.json"

name_df <- tribble(
  ~station_code, ~city_name, ~lat, ~long, ~station_name,
  "009151", "Perth", -31.96, 115.79, "Subiaco Wastewater Treatment Plant",
  "023011", "Adelaide", -34.92, 138.6, "North Adelaide",
  "040383", "Brisbane", -27.51, 153.05, "Greenslopes Private Hospital",
  "040913", "Brisbane", -27.48, 153.04, "Brisbane",
  "066062", "Sydney", -33.86, 151.21, "Observatory Hill",
  "070351", "Canberra", -35.31, 149.2, "Canberra Airport",
  "086232", "Melbourne", -37.83, 144.98, "Melbourne Botanical Gardens"
)
```


Sidebar {.sidebar}
======================================================================
Use the _Explore rainfall_ tab to explore the the rainfall by the major cities in Australia. The blue line on the time series plot shows long term trend of monthly rainfall. Click and drag on the time series plot to highlight the date and the corresponding monthly average rainfall. Search the data table to view date and the corresponding monthly rainfall. 

Use the _Rain fall_ tab to explore the the temperature by the major cities in Australia. The red and yellow line shows the long term of minimum and maximum temperature trend. Click and drag on the time series plot to highlight the daily temperature.Search the data table to view the date and corresponding daily temperature

Application Authors: Alison, Kate, Keke

Data source: (http://www.bom.gov.au/?ref=logo) Bureau of Metrology, Australia


Rain fall
=====================================================================


Column {.tabset}
----------------------------------------------------------------------

### Rain fall plot

```{r}
clean_rain_df_3 <- rainfall %>%
  filter(year > 2009) %>%
  mutate(year_month = paste(year, month, sep = "-")) %>%
  mutate(year_month = ym(year_month)) %>%
  group_by(station_code, city_name, lat, long, station_name, year_month) %>%
  dplyr::summarize(monthly_rain = sum(rainfall)) 

rain_plot_monthly <- clean_rain_df_3 %>%
  ggplot() +
  geom_point(aes(x = year_month, y = monthly_rain), 
             color = "blue", alpha = .8, size = .1) + 
  geom_smooth(aes(x = year_month, y = monthly_rain), color = "black") + facet_wrap(vars(city_name)) +
  theme_classic() + ylab("Total Monthly Rainfall (cm)") + 
  xlab("Date (year)")

ggplotly(rain_plot_monthly)
```

### Rainfall data table 

```{r}
clean_rain_df_3 %>%
  rename(`station code` = station_code,
         `city name` = city_name,
         `station name` = station_name,
          year = year_month,
         longitude = long,
         latitude = lat,
           `monthly rain` = monthly_rain) %>%
datatable()
```

Temperature
============================================================================

Column {.tabset}
----------------------------------------------------------------------

### Temperature plots

```{r}
clean_temp_df <- temperature %>%
group_by(city_name, temp_type, site_name, date) %>%
dplyr::summarize(average_temp = mean(temperature, na.rm = TRUE)) %>%
pivot_wider(names_from = temp_type, values_from = average_temp)

clean_temp_df_all <- temperature %>% 
  pivot_wider(names_from = temp_type, values_from = temperature)
clean_temp_df_all$year <- substr(clean_temp_df_all$date, 1, 4)
clean_temp_df_10 <- clean_temp_df_all %>%
  filter(year > 2009)

temp_plot_10 <- clean_temp_df_10 %>%
  ggplot() +
  geom_point(aes(x = date, y = max), size = .1, alpha = .05, color = "red") +
  geom_point(aes(x = date, y = min), size = .1, alpha = .05, color = "orange") +
  geom_smooth(aes(date, max), color = "red") +
  geom_smooth(aes(date, min), color = "orange") +
  facet_wrap(vars(city_name)) +
  theme_classic() + ylab("Max and Min Temp (C)") + xlab("Date (year)")

ggplotly(temp_plot_10)
```

### Temperature data table 
```{r}
clean_temp_df %>%
  
  rename(`site name` = site_name,
         `city name` = city_name) %>%
  
datatable()
```



