---
title: "Kate_Code"
author: "Kate Bazany"
date: "2022-11-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, Kate}
library(tidytuesdayR)
```


#Adding Wildfire Data

```{r, Kate}
# Get the Data

rainfall <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-07/rainfall.csv')
temperature <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-07/temperature.csv')

# IF YOU USE THIS DATA PLEASE BE CAUTIOUS WITH INTERPRETATION
nasa_fire <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-07/MODIS_C6_Australia_and_New_Zealand_7d.csv')

# For JSON File of fires
url <- "http://www.rfs.nsw.gov.au/feeds/majorIncidents.json"

aus_fires <- sf::st_read(url)

# Or read in with tidytuesdayR package (https://github.com/thebioengineer/tidytuesdayR)

# Either ISO-8601 date or year/week works!

# Install via devtools::install_github("thebioengineer/tidytuesdayR")

tuesdata <- tidytuesdayR::tt_load('2020-01-07') 
tuesdata <- tidytuesdayR::tt_load(2020, week = 2)

rainfall <- tuesdata$rainfall
```

Example Plot

```{r}
# Mapping NSW Current Incidents in R -------------------------------------------

library(sf)
library(mapview)
library(janitor)
library(tidyverse)

#' Current Incidents Feed (GeoJSON)
#' This feed contains a list of current incidents from the NSW RFS, 
#' and includes location data and Major Fire Update summary information where available. 
#' Click through from the feed to the NSW RFS website for full details of the update. 
#' GeoJSON is a lightweight data standard that has emerged to support the sharing of 
#' information with location or geospatial data. 
#' It is widely supported by modern applications and mobile devices.

url <- "http://www.rfs.nsw.gov.au/feeds/majorIncidents.json"

fires <- st_read(url)

fires

mapview(fires)

#' Hacky way to get rid of points within geometry collections
fire_poly <- fires %>% 
  st_buffer(dist = 0) %>% 
  st_union(by_feature = TRUE)

mapview(fire_poly)

fires %>% 
  mutate(pubdate = as.character(pubDate),
         pubdate = as.Date(pubdate))
```

Cleaning Data

```{r}
library(tidyverse)
library(here)

read_file_list <- list.files(here::here("2020", "2020-01-07")) %>% 
  .[str_detect(., "tmax|tmin")]

read_clean_temp_data <- function(file_name){
  
  temp_descrip <- if_else(str_detect(file_name, "min"), "min", "max")
  
  read_csv(here::here("2020", "2020-01-07", file_name)) %>% 
    janitor::clean_names() %>% 
    fill(site_name, site_number) %>% 
    filter(!is.na(date)) %>% 
    rename(temperature = contains("temp")) %>% 
    mutate(temp_type = temp_descrip)  %>% 
    mutate(city_name = word(site_name, 1)) %>%  
    select(city_name, date, temperature, temp_type, site_name)
  
}


# Get Clean Temp Data -----------------------------------------------------

clean_df <- read_file_list %>% 
  map(read_clean_temp_data) %>% 
  bind_rows()

write_csv(clean_df, here::here("2020", "2020-01-07","temperature.csv"))


##### Prep for Clean Rain Data
#####
#####

name_df <- tribble(
  ~station_code, ~city_name, ~lat, ~long, ~station_name,
  "009151", "Perth", -31.96, 115.79, "Subiaco Wastewater Treatment Plant",
  "023011", "Adelaide", -34.92, 138.6, "North Adelaide",
  "040383", "Brisbane", -27.51, 153.05, "Greenslopes Private Hospital",
  "040913", "Brisbane", -27.48, 153.04, "Brisbane",
  "066062", "Sydney", -33.86, 151.21, "Observatory Hill",
  "070351", "Canberra", -35.31, 149.2, "Canberra Airport",
  "086232", "Melbourne", -37.83, 144.98, "Melbourne Botanical Gardens"
)

write_csv(name_df, "station_names_locations.csv")

read_precip_list <- list.files(here::here("2020", "2020-01-07")) %>% 
  .[str_detect(., "IDCJ")]

read_clean_precip_data <- function(file_name){
  
  read_csv(here::here("2020", "2020-01-07", file_name)) %>% 
    janitor::clean_names() %>% 
    select("station_code" = bureau_of_meteorology_station_number,
           year, month, day,
           "rainfall"= rainfall_amount_millimetres, 
           "period" = period_over_which_rainfall_was_measured_days,
           quality) %>% 
    mutate(station_code = as.character(station_code)) %>% 
    left_join(name_df, by = "station_code") %>% 
    select(station_code, city_name, everything())
  
}


# Get Clean Rain Data -----------------------------------------------------

clean_rain_df <- read_precip_list %>% 
  map(read_clean_precip_data) %>% 
  bind_rows()

write_csv(clean_rain_df, here::here("2020", "2020-01-07","rainfall.csv"))
```

Trying an alternative to getting rainfall and temperature data

Rainfall

```{r}
library(lubridate)

rainfall$rainfall <- as.integer(rainfall$rainfall)

rainfall

clean_rain_df <- rainfall %>%
  group_by(station_code, city_name, lat, long, station_name, year) %>%
  summarize(yearly_rain = sum(rainfall)) 

write_csv(clean_rain_df, "clean_rain.csv")

clean_rain_df_2 <- rainfall %>%
  mutate(date = paste(year, month, day, sep = "-")) %>%
  mutate(date_date = ymd(date))

clean_rain_df_2
write_csv(clean_rain_df_2, "clean_rain_df_d.csv")

clean_rain_df_all <- rainfall %>%
  mutate(date = paste(year, month, day, sep = "-"))
clean_rain_df_all$date <- ymd(clean_rain_df_all$date)
clean_rain_df_all
```



Temperature

```{r}
library(tidyverse)

temperature$temperature <- as.integer(temperature$temperature)

temperature

clean_temp_df <- temperature %>%
  group_by(city_name, temp_type, site_name, date) %>%
  summarize(average_temp = mean(temperature, na.rm = TRUE)) %>%
  pivot_wider(names_from = temp_type, values_from = average_temp)

write_csv(clean_temp_df, "clean_temp.csv")

clean_temp_df_all <- temperature %>% 
  pivot_wider(names_from = temp_type, values_from = temperature)
clean_temp_df_all
```

dygraphs package?
cross talk?
popups?

map color code by type of fire?
size by brightness?

rain_plot <- clean_rain_df_2 %>%
  ggplot() +
  geom_point(aes(x = date_date, y = rainfall)) +
  facet_wrap(vars(station_name))
  
  
rain_plot

```{r}
clean_rain_df_2
clean_temp_df

class(clean_rain_df$year)
clean_rain_df$year <- as.integer(clean_rain_df$year)

rain_plot <- clean_rain_df %>%
  ggplot() +
  geom_smooth(aes(x = year, y = yearly_rain)) +
  facet_wrap(vars(city_name)) +
  theme_classic() + ylab("Total Yearly Rain (cm)") + ylim(0, 2000)

rain_plot

library(plotly)

ggplotly(rain_plot)
```


```{r}
class(clean_temp_df)

temp_plot <- clean_temp_df %>%
  ggplot() +
  geom_smooth(aes(x = date, y = max), color = "red") +
  geom_smooth(aes(x = date, y = min), color = "orange") +
  facet_wrap(vars(city_name)) +
  theme_classic() + ylab("Max and Min Temp (C)") + xlab("Year")

temp_plot

ggplotly(temp_plot)
```


Add a zoomed in plot of the last 20 years of rainfall.

(consider transforming (add .1 to get rid of NA values), add back daily data)

```{r}
clean_rain_df_sub <- clean_rain_df %>%
  filter(year > 2000) %>%
#  filter(city_name != "Adelaide") %>%
  mutate(rain_trans = yearly_rain + 1)

rain_plot_sub <- clean_rain_df_sub %>%
  ggplot() + 
  geom_vline(xintercept = 2019.94, color = "red", size = 1.2, alpha = .8) +
  geom_smooth(aes(x = year, y = rain_trans)) +
  facet_wrap(vars(city_name)) +
  theme_classic() + ylab("Total Yearly Rain (cm)") + ylim(0, 2000)

rain_plot_sub

ggplotly(rain_plot_sub)
```

Zoomed in Plot of temperature

```{r}
temperature
rainfall
```

Zoomed in Plot of Rainfall

```{r}

```


Full temperature plot

```{r}
library(ggplot2)
library(dplyr)
temp_plot_full <- clean_temp_df_all %>%
  ggplot() +
#  geom_vline(xintercept = 2020, color = "red", size = 1.2, alpha = .8) +
  geom_point(aes(x = date, y = max), size = .1, alpha = .05, color = "red") +
  geom_point(aes(x = date, y = min), size = .1, alpha = .05, color = "orange") +
#  geom_smooth(aes(x = date, y = max), color = "red") +
#  geom_smooth(aes(x = date, y = min), color = "orange") +
  facet_wrap(vars(city_name)) +
  theme_classic() + ylab("Max and Min Temp (C)") + xlab("Year")

temp_plot_full
```

Full Rainfall Plot

```{r}
clean_rain_df_all
rain_plot_full <- clean_rain_df_all %>%
  ggplot() +
  geom_point(aes(x = date, y = rainfall), color = "blue", alpha = .05, size = .1) +
  facet_wrap(vars(city_name)) +
  theme_classic() + ylab("Rainfall (cm)")
rain_plot_full
```


