---
title: "R_Group_Project"
author: "Alison Barbee"
date: "2022-12-14"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



#…or create a new repository on the command line
echo "# R_Group_Project" >> README.md
git init
git add README.md
git commit -m "first commit"
git branch -M main
git remote add origin git@github.com:alisonbarbee/R_Group_Project.git
git push -u origin main



#…or push an existing repository from the command line
git remote add origin git@github.com:alisonbarbee/R_Group_Project.git
git branch -M main
git push -u origin main


Adding Wildfire Data

```{r}
# Get the Data

rainfall <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-07/rainfall.csv')
temperature <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-07/temperature.csv')

# IF YOU USE THIS DATA PLEASE BE CAUTIOUS WITH INTERPRETATION
nasa_fire <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-07/MODIS_C6_Australia_and_New_Zealand_7d.csv')

# For JSON File of fires
url <- "http://www.rfs.nsw.gov.au/feeds/majorIncidents.json"

aus_fires <- sf::st_read(url)

# Or read in with tidytuesdayR package (https://github.com/thebioengineer/tidytuesdayR)

# Either ISO-8601 date or year/week works!

# Install via devtools::install_github("thebioengineer/tidytuesdayR")

tuesdata <- tidytuesdayR::tt_load('2020-01-07') 
tuesdata <- tidytuesdayR::tt_load(2020, week = 2)


rainfall <- tuesdata$rainfall
```



